\section*{Abstract}

Policy microsimulation models rely on imputation to attach variables one
survey observes to records from another, and the imputed file is then
reweighted --- calibrated to administrative totals, filtered to
subpopulations, stressed under reform scenarios. Common imputation practice
fits models to survey records while ignoring their design weights, which
biases imputed conditionals toward the sample rather than the population and
plants extreme low-weight donor values that a later reweighting can amplify.
We present the imputation operator of \populace{}, \policyengine{}'s
open-source microdata stack: a regime-gated, sequentially-chained,
weighted-bootstrap quantile-regression-forest estimator whose interface makes
an unweighted fit impossible to request by accident. To evaluate it we formalize a
\emph{population-view harness}: one latent population observed through survey
views, with any candidate weighted file scored by projecting it through each
view against that survey's holdout --- a strictly proper joint score, a
coverage axis invariant to reweighting of the candidate, a classifier
two-sample test, and an uncapped tail block on the imputed columns, which we
add after showing that subsampled joint geometry certifies files whose top
percentile is wrong by a factor of two. Under this harness and paired repeated holdouts we benchmark
the estimator against standard survey-imputation methods --- unweighted and
weighted quantile regression forests, ordinary least squares, linear quantile
regression, and nearest-neighbour hot-deck statistical matching --- on
within-\scf{} wealth, \scf{}-onto-\cps{} population synthesis, zero-inflated
\cps{} income components, and six cross-domain datasets. Ablations attribute performance to each design choice,
and the experiment runs at two scopes: a minimal six-predictor instantiation
and a populace-scale one (ten shared predictors, a four-component chained
balance sheet, three pooled receiver vintages). Where the survey design is
informative, weighting dominates: the identical estimator fit unweighted is
six times worse on net-worth marginal fit within the donor survey, and on the
population view its imputed net-worth 99th percentile is 1.9 to 2.3 times the
holdout's --- an inflation the geometry metrics tie on and only the tail
block detects. Regime gating carries zero-inflated components (0.6 versus 4.4
percentage points of dividend zero-share error against the same forest
ungated, and 25--84 points against linear and Gaussian baselines). Chaining
is a joint property that binds at scale: with two targets it is
indistinguishable, while on the four-component balance sheet fitting targets
independently costs 5.8 points of two-sample AUC --- the classifier detects
balance sheets that do not add up. Hot-deck matching is a genuine near-peer
on the population view; marginal metrics alone rank an unconditional weighted
draw at or near the top throughout, and at populace scale the classifier
separates it almost perfectly (AUC 0.97 versus the candidate's 0.76). A
closed-form reweighting stress test shows every tail-capable method matching
the held-out truth's own concentration while Gaussian-residual imputation
understates it by factors of 7 to 11 --- missing tails, not robustness. Wealth imputation is what
makes asset-tested programs such as Supplemental Security Income simulable on
the \cps{} at all, and the population-view results identify which methods
preserve the demographics--wealth joint that such simulation depends on. The
estimator is available as \populacefit{}, installable independently of the
rest of the stack.

\vspace{1em}
\noindent\textbf{Keywords:} microsimulation, imputation, survey weights,
quantile regression forests, statistical matching
