\section{Data and tasks}
\label{sec:data}

Four task families span the settings the estimator is used in. Where a task
imputes within one survey, the protocol of Section~\ref{sec:protocol} applies:
80/20 donor/receiver holdouts repeated over ten paired seeds. The
population-view experiment additionally projects candidates onto a second
survey. Every table in Section~\ref{sec:results} regenerates from the
repository's committed run configurations; row caps and skipped cells are
recorded in the run manifests.

\subsection{Task 1: wealth within the SCF}
\label{sec:task-scf}

The Survey of Consumer Finances \citep{fed2023scf} observes household wealth
in detail. From the 2022 summary extract (first implicate, survey weights) we
impute \texttt{debt} --- zero-inflated and nonnegative --- and
\texttt{networth} --- sign-mixed, with a heavy right tail --- from age, sex,
education class, marital status, children, total income, and wage income.
The two targets are chosen deliberately: their sign structures exercise the
regime gates, and their strong mutual dependence exercises chaining.

\subsection{Task 2: the SCF\texorpdfstring{$\to$}{->}CPS population view}
\label{sec:task-harness}

Each method fits wealth conditionals on an SCF donor split and imputes them
onto real \cps{} ASEC households sharing the predictor set --- households
built from the Census Bureau's ASEC public-use files \citep{census2025}, read
directly from \texttt{census.gov} so that no task input flows through any
PolicyEngine processing. The experiment runs in two profiles. The
\emph{minimal} profile --- six shared predictors, the two wealth targets of
Section~\ref{sec:task-scf}, a single receiver vintage (ASEC 2025) --- is the
controlled instantiation in which every mechanism is legible. The
\emph{populace-scale} profile mirrors the production stack's shape: the
shared predictor set widens to everything the two surveys genuinely share
(ten variables, adding education class, race, homeownership, and labor-force
status, each mapping verified against both files' codebooks), the receiver
pools three ASEC vintages (2023--2025, weights scaled to one household
population, 168,852 households) as the production support spine does, and
the targets widen to a four-component chained balance sheet --- financial
assets, nonfinancial assets, debt, then net worth --- whose accounting
identity $\text{networth} = \text{fin} + \text{nfin} - \text{debt}$ holds
exactly in the extract, so chaining's contribution is directly
interpretable. The candidate population --- \cps{} demographics plus imputed wealth,
under \cps{} household weights --- is then scored through the harness of
Section~\ref{sec:harness} against the \emph{held-out} SCF: the joint of
shared predictors and wealth, under SCF weights. The
\texttt{scf\_sample\_reference} rows score the donor split itself against the
holdout, anchoring the sampling-noise floor. This is the paper's direct test
of whether conditional-distribution methods preserve joint structure that
matching's conditional-independence gluing loses \citep{rassler2002statistical,
dorazio2006statistical}, measured where it matters: on a receiver file the
donor survey never saw. Because the demographic block of every candidate is
the same real \cps{} sample, differences between methods isolate the wealth
block and its coupling to demographics; the shared demographic offset between
the two surveys is common to all methods and bounded below by the reference
floor.

\subsection{Task 3: zero-inflated income components in the CPS}
\label{sec:task-cps}

From the Census Bureau's ASEC 2025 public-use person file (income reference
year 2024), read directly from \texttt{census.gov} --- never a processed or
enhanced artifact, which would embed the production imputations this paper is
about --- we impute interest and dividend income for adults from age, sex,
and employment income, under the ASEC person supplement weights. (Survey
files publish per-person design weights; \populace{}'s production convention
deliberately carries calibrated \emph{household} weights onto persons, since
its calibration operates at household grain. The task fixes the survey's own
published weights, and because every method shares whatever weights the task
fixes, the method comparison does not turn on this choice.) Both components
are zero for most adults (41.7 and 85.2 percent, respectively) with heavy
positive tails --- the regime surface where gating and weighting matter
most. Public-use ASEC amounts are topcoded, which truncates the
extreme tail relative to an administrative or SCF-style source; the fragility
comparisons of Section~\ref{sec:results-cps} are therefore conservative. This task also carries
the reweighting stress test of Section~\ref{sec:metrics}: prior
\policyengine{} production experience motivates it, where an unweighted fit
broadcast rare high-value donor records across a receiver file as
near-zero-weight point masses that later reweighting amplified. All fragility
numbers reported here are measured under this paper's protocol.

\subsection{Task 4: cross-dataset transfer}
\label{sec:task-openml}

To test behavior beyond economic microdata, we run the method surface over
the six OpenML AutoML Benchmark regression datasets used in the
\microimpute{} manuscript's appendix \citep{vanschoren2014openml} ---
\texttt{space\_ga}, \texttt{elevators}, \texttt{brazilian\_houses},
\texttt{onlinenewspopularity}, \texttt{abalone}, and \texttt{house\_sales}
--- treating each dataset's target as the imputed column under uniform
weights.

Full multi-view synthesis --- one candidate scored simultaneously against
\cps{}, \sipp{} \citep{census2023sipp}, and \psid{} \citep{psid2023} views ---
is the production release-gate use of the harness and is left to future work;
\psid{} access requires registration, and the two-view experiment of
Section~\ref{sec:task-harness} already identifies the joint-structure
differences the paper is about.