\section{Discussion}
\label{sec:discussion}

The results attribute the estimator's behavior to its design choices with a
regularity worth stating plainly. \emph{Weighting is the dominant choice
wherever the design is informative}: on the \scf{}, whose list sample
oversamples wealthy households, removing the weighted bootstrap costs a
factor of six on net-worth marginal fit within the donor survey and inflates
the imputed 99th percentile on the population view by a factor of 1.9 to 2.3
--- larger than the gap between the candidate and any competing method. On
the \cps{} components task, whose person weights carry little outcome
information, weighting is approximately free. An estimator cannot know in
advance which regime it is in, which is the argument for weighting by
construction rather than by option. \emph{Gating carries the zero-inflation
surface} (0.6 versus 4.4 percentage points of dividend zero-share error
against the same forest ungated, and 5.6 AUC points on the populace-scale
population view). \emph{Chaining is a joint property that binds at scale}:
invisible with two targets, it is worth 5.8 AUC points when the four imputed
components must satisfy a balance-sheet identity --- and only the omnibus
classifier sees the violation, because a balance sheet that does not add up
has perfectly reasonable marginals. None of the three choices hurts where it
does not help.

The evaluation findings are as load-bearing as the method findings, and one
of them corrected this paper's own harness. Marginal metrics saturate: with
weak predictors, an unconditional weighted draw from the donor marginal ties
or beats every conditional method on pinball loss and Wasserstein distance
--- on the OpenML suite it attains the best mean rank --- while carrying no
conditional structure at all; at populace scale the classifier separates it
almost perfectly (AUC 0.968 against 0.761). But the joint metrics have their
own blind spot: capped pairwise geometry tied the unweighted ablation with
the candidate while its imputed net-worth 99th percentile was double the
holdout's, because the discrepant mass is a sliver of standardized pairwise
space. We caught this by comparing weighted quantiles directly, added the
uncapped tail block to the harness, and report it as a general lesson:
sample-geometry evaluations of heavy-tailed economic microdata need an
explicit tail axis, with a sampling floor to calibrate its noise (the floor's
own q99 ratio wobbles to 1.10). The fragility diagnostic adds the
complementary caution --- the target is the truth's own tail exposure, not
minimal exposure, and the one method that minimizes it (OLS, at factors of 7
to 11 below truth) does so by failing to generate realistic tails at all.

Several limitations bound the claims. First, the population-view experiment
runs one view pair: the common \cps{}--\scf{} offset (differing income
concepts, ASEC topcoding, and income reference years spanning 2021 to 2024
across the pooled vintages) is shared by all methods but bounds how close any
candidate can come to the floor; a multi-view instantiation (adding \sipp{}
and \psid{}) would tighten the identification and is the production
release-gate use of the harness. Second, the joint-geometry blocks still run
on weight-proportional subsamples (2,048 points per side); the tail block and
the C2ST, which use all rows, are the axes least affected. Third, no
conditional method gets the extreme tail right: forest draws do not
extrapolate beyond training support, so imputed q99 ratios sit at 0.70--0.81
for the candidate and its forest relatives, where donation-based draws sit
near 1 --- a genuine advantage of hot-deck matching that the tail block keeps
visible, and a caution against reading the candidate's wins as uniform.
Baseline quantile models are additionally converted to samplers through a
99-level grid, truncating draws beyond the 1st and 99th conditional
percentiles. Fourth, the \microimpute{} forest's internal randomness is not
seedable through its public constructor, so its paired-seed dispersion
includes forest noise the candidate's does not. Fifth, comparisons hold
shared forest backends and default hyperparameters fixed rather than tuning
each method per task; the ablations are exact controls, the cross-family
comparisons are defaults-versus-defaults. Finally, we evaluate single
imputations distributionally and do not propagate imputation uncertainty in
the multiple-imputation sense \citep{rubin1987multiple}; for the population
statistics microsimulation consumes, the distributional view is the relevant
one, but inference from imputed files would require the fuller treatment.
