\section{Introduction}

Microsimulation models of tax and transfer policy require individual records
that span domains no single survey covers
\citep{bourguignon2006microsimulation, sutherland2013euromod}. The Current
Population Survey observes employment and income but not wealth; the Survey of
Consumer Finances observes wealth in detail but lacks the \cps{}'s coverage
and program detail; administrative tax records observe income components
precisely but only for filers, and with narrow demographics. Model builders
bridge these gaps with imputation: fit a conditional model of the missing
variables on a donor survey, then draw values for the receiver's records
\citep{little2002statistical, dorazio2006statistical}.

Two properties of this setting distinguish it from generic missing-data
problems. First, both files are \emph{weighted samples}: each record stands in
for a different number of population units, and the quantity of interest is a
population distribution, not a sample one. An imputation model fit to
unweighted records recovers the sample conditional, which differs from the
population conditional exactly when the design (or a prior calibration)
correlates weights with outcomes. Second, the imputed file is not an endpoint.
It is \emph{reweighted downstream} --- calibrated to administrative margins
\citep{deville1992calibration, hainmueller2012entropy}, filtered to
subpopulations, stressed under reforms --- so an imputation that looks
acceptable under the weights it shipped with can fail under weights it never
saw. A rare donor record carrying an extreme value at near-zero weight is
invisible in the shipped aggregates; a later reweighting that assigns it mass
turns it into a distortion of the population total. We refer to such records
as landmines, and we measure exposure to them directly with a reweighting
stress test (\todo{fragility results reference}).

This paper presents and evaluates the imputation operator of \populace{},
\policyengine{}'s open-source microdata stack \citep{populace2026}. The
estimator combines three design choices, each motivated by one of the failure
modes above: a \emph{weighted bootstrap} that materializes survey weights into
the training data of each forest, so the learned conditionals are population
conditionals; \emph{regime gates} that classify each record's sign class
(negative, zero, positive) before magnitudes are modeled, so zero-inflated and
sign-mixed variables --- ubiquitous in economic microdata
\citep{mullahy1986specification, lambert1992zero} --- retain their mass
structure; and \emph{sequential chaining} across imputed variables, so joint
structure among targets survives the imputation
\citep{raghunathan2003multiple, vanbuuren2011mice}. The interface enforces the
first choice: fits are weighted by construction, and an unweighted fit must be
requested explicitly.

We make three contributions:

\begin{enumerate}
\item \textbf{A weight-aware imputation estimator, available standalone.}
We describe the estimator and its interface contract, and release it as
\populacefit{}, installable and usable on plain data frames independently of
the rest of the \populace{} stack.
\item \textbf{The population-view harness.} We formalize the evaluation
question as one latent population observed through survey \emph{views} ---
variable subsets, designs, measurement idioms --- and score any candidate
weighted file by projecting it through each view against that survey's
holdout, using a strictly proper joint score \citep{gneiting2007strictly,
szekely2013energy}, a support-based coverage axis that is invariant to
reweighting of the candidate \citep{naeem2020reliable}, and a weighted
classifier two-sample test. The harness is generator-agnostic and
non-self-referential (holdouts appear nowhere upstream), and it cleanly
separates what record generation must get right from what calibration can
later change.
\item \textbf{A benchmark against standard practice with attribution, plus a
reweighting stress test.} We compare the estimator against unweighted and
weighted quantile regression forests \citep{meinshausen2006quantile},
ordinary least squares imputation \citep{vonhippel2007should}, linear
quantile regression \citep{koenker1978regression}, and nearest-neighbour
hot-deck statistical matching \citep{andridge2010review,
dorazio2021statistical}, under a paired, repeated holdout protocol. Ablations
knock out one design choice at a time, attributing performance to weighting,
gating, and chaining separately, and a closed-form fragility diagnostic
measures each method's worst-case single-record exposure under bounded
reweightings of the imputed file --- the landmine failure mode.
\end{enumerate}

The policy stakes are concrete. Supplemental Security Income conditions
eligibility on countable resources \citep{ssa2024ssi}, so simulating its
baseline caseload on the \cps{} --- or reforms such as the SSI Savings Penalty
Elimination Act, which would raise the resource limits
\citep{ssi_spea_2025} --- is impossible without imputed wealth, and the
composition of the simulated caseload is exactly a functional of the imputed
demographics--wealth joint. The population-view experiment of
Section~\ref{sec:task-harness} scores that joint directly: it asks whether
households assembled from \cps{} demographics and imputed wealth are
distributionally exchangeable with households the \scf{} actually observed.

The rest of the paper proceeds as follows. Section~2 situates the estimator in
the survey-imputation literature. Section~3 specifies the estimator, the
ablation design, the metrics, and the protocol. Section~4 describes the four
benchmark tasks and their data. Section~5 reports results, Section~6
discusses limitations, and Section~7 concludes.
