\section{Methodology}

\subsection{The estimator}
\label{sec:estimator}

\populacefit{} models $P(\text{targets} \mid \text{predictors})$ with a
quantile-regression-forest estimator \citep{meinshausen2006quantile,
zillow2024quantile} wrapped in three mechanisms.

\paragraph{Weighted bootstrap.} Random forests cannot honor a per-record
\texttt{sample\_weight} in their predictive distribution: a fully grown leaf
holds individual training rows, and weighting the impurity criterion does not
change which values a quantile query reads out. \populacefit{} therefore
materializes weights into the data: before each forest is grown, training rows
are resampled with replacement with probability proportional to weight, so
leaf distributions reflect the weighted population. Draws from the fitted
forest then sample the \emph{weighted} conditional.

\paragraph{Regime gates.} A numeric target's sign support --- which of
$\{\text{negative}, \text{zero}, \text{positive}\}$ occur --- defines its
regime. A single regressor over a zero-inflated or sign-mixed target either
drops a tail or interpolates across the empty gap at zero; hurdle-style
two-part models are the classical remedy \citep{mullahy1986specification,
lambert1992zero}. \populacefit{} fits a classifier that gates each row into a
sign class and a separate magnitude forest within each nonzero sign. Regime
detection is structural (it reads the unweighted support, since which signs
\emph{exist} is a fact about the variable, not the weighting), while the gate
itself is weighted directly through its \texttt{sample\_weight} --- not by
resampling, which would delete a rare sign class whose total weight share is
negligible. Draws sample the gate's predicted class probabilities rather than
taking the modal class, preserving the zero mass.

\paragraph{Sequential chaining.} Targets are imputed in order, each
conditioning on the predictors plus the targets already drawn
\citep{raghunathan2003multiple, vanbuuren2011mice}, so dependence among
imputed variables survives.

\paragraph{Interface contract.} Fitting over a \populace{} frame reads the
frame's typed survey weights by default; fitting over a plain data frame ---
the standalone path --- requires the weights explicitly (a weight column name,
a weight vector, or the literal \texttt{"none"}), and omitting them is an
error rather than a silent unweighted fit. The contract makes the
weight-blindness failure mode of Section~2.3 unrepresentable by accident.

\subsection{Ablations and baselines}
\label{sec:ablations}

Each design choice is evaluated by knocking it out with everything else held
fixed: the full estimator versus itself with \texttt{weights="none"} (the
weighted bootstrap's contribution), versus fitting each target independently
(chaining's contribution), and versus a single ungated forest at matched
hyperparameters (gating's contribution). Baselines are standard practice:
quantile-regression-forest, ordinary-least-squares, and linear
quantile-regression imputers from \microimpute{}
\citep{policyengine2025microimpute}, and nearest-neighbour-distance hot-deck
statistical matching from \pystatmatch{} \citep{dorazio2021statistical}. The
candidate and the plain-forest baselines share the same underlying forest
implementation \citep{zillow2024quantile}, so their differences isolate the
design choices rather than implementation quality.

\subsection{The population-view harness}
\label{sec:harness}

Per-variable metrics cannot adjudicate the question this paper cares about,
because the methods differ most in how they treat the \emph{joint}. We
therefore evaluate within a framework we call the population-view harness. One
latent population distribution produces individuals; each survey $s$ is a
\emph{view} $V_s$ of it --- a variable subset, a sampling design, a
measurement idiom --- and the survey's holdout is a weighted sample from
$V_s(P)$ \citep{lohr2017combining}. A candidate weighted file $Q$ (any
generator's) is scored by projecting it through each view and comparing
$V_s(Q)$ to that view's holdout \emph{in the view's own variable space}.
Cross-survey consistency is never required --- the candidate is only asked to
explain each view in its own idiom --- which keeps the harness usable even
though surveys disagree with one another about the population they share.

Each view carries four complementary blocks. The \emph{weighted energy
distance} \citep{szekely2013energy} is the sample form of a strictly proper
scoring rule \citep{gneiting2007strictly}: the true distribution uniquely
minimizes its expectation, so a candidate hedged toward modal households
scores strictly worse than one matching the full distribution. \emph{\prdc{}
coverage} \citep{naeem2020reliable} --- the weighted fraction of holdout
points with a candidate point inside their $k$-nearest-neighbour radius ---
is the explicit anti-collapse axis; because it depends on the candidate only
through its support, it is invariant to any reweighting of the candidate,
making it the calibration-blind block of the scorecard. A \emph{weighted
classifier two-sample test} (cross-validated AUC of a gradient-boosted
classifier at equal class mass; $0.5$ is indistinguishable) is the omnibus
check. The fourth block exists because we caught the first three missing a
real failure: pairwise sample-geometry metrics run on capped subsamples, and
a candidate whose imputed wealth 99th percentile is double the holdout's can
tie on energy distance because the discrepant mass is a sliver of
standardized pairwise space. Economic variables live in heavy right tails, so
each \emph{imputed} column also carries a \emph{tail block}, computed on the
full weighted samples with no cap: its weighted Wasserstein-1 distance to the
holdout (scaled by the holdout's weighted standard deviation) and its
weighted q90 and q99 ratios (candidate over holdout; 1 is a perfect tail).
Holdouts are used nowhere upstream, so the harness is a non-self-referential
test surface; a sampling-noise floor --- the donor split itself scored as a
candidate --- anchors what "as good as another sample of the survey" means on
each axis, including how much a tail ratio wobbles under sampling noise
alone.

\subsection{Marginal and stress metrics}
\label{sec:metrics}

All metrics are computed under survey weights, on held-out receiver rows.
\emph{Weighted pinball loss}, averaged over a decile grid, scores conditional
distributional calibration \citep{koenker2005quantile}. \emph{Weighted
Wasserstein-1 distance} to the weighted donor distribution scores marginal
fit. \emph{Zero-share error} scores preservation of the exact-zero mass of
zero-inflated targets. Finally, \emph{reweighting fragility} formalizes the
landmine diagnostic: over the family of bounded multiplicative reweightings
$w_i \mapsto m_i w_i$ with $m_i \in [1/\kappa, \kappa]$ (the hard weight-ratio
bounds production calibration guards use), the worst-case single-record share
of a target aggregate has the closed form
$\kappa^2 c^* / (\kappa^2 c^* + S - c^*)$, where $c^*$ is the largest baseline
contribution $w_i |a_i|$ and $S$ their sum. We report it at $\kappa = 5$ for
every method next to the held-out truth's own fragility --- the exposure a
method should not exceed.

\subsection{Protocol}
\label{sec:protocol}

Every task uses an 80/20 donor/receiver holdout repeated over ten seeds, with
splits paired across methods: the split is a pure function of the seed, so
every method fits and is scored on identical partitions. Task inputs are
frozen once and pinned by hash; only the method varies within a sweep. All
results tables are generated from the sweep artifacts by the repository's
command line; no result is hand-entered. Where a method is unavailable or a
metric undefined for a task, the cell is reported as absent rather than
omitted.
